---
title: "Free Plan Usage"
output: github_document
---

## Motivation
The purpose of this analysis is to get a better understanding of how users on the Individual plan use Buffer, in order to better inform decisions around the plan limits.

### How often do active users post?
To answer this question, we'll need to gather users, their join date, their last update date (if it was in the past 28 days, they are _currently_ active), and the number of updates they have shared with Buffer

```{r include = FALSE}
library(buffer); library(dplyr); library(ggplot2)
```

We'll use the data in [this look](https://looker.buffer.com/looks/3919) to inform our analysis.

```{r}
# Get users from look
users <- get_look(3919)
```

Cool, there are 243k individual users that have shared at least one Buffer update in the past 28 days. Let's visualize the number of updates per day for each of them.

```{r include = FALSE}
# Rename columns
colnames(users) <- c('user_id', 'created_at', 'updates', 'days')

# Set dates as date objects
users$created_at <- as.Date(users$created_at)

# Calculate days since joined
users <- users %>%
  mutate(days_since_join = as.numeric(Sys.Date() - created_at))

# Calculate updates per day
users <- users %>%
  mutate(updates_per_day = updates / days_since_join,
         updates_per_update_day = updates / days)
```

```{r echo = FALSE}
ggplot(users, aes(x = updates_per_day)) + 
  geom_density() +
  scale_x_continuous(limits = c(0, 100))
```

So most users post very few times per day, but a very small number of users post many times per day. Let's break it down into buckets.

```{r echo = FALSE}
users <- users %>%
  mutate(updates_rounded = round(updates_per_day, 0))

# Group by updates
by_updates <- users %>%
  group_by(updates_rounded) %>%
  summarise(users = n_distinct(user_id))

ggplot(filter(by_updates, updates_rounded <= 10), aes(x = as.factor(updates_rounded), y = users)) +
  geom_bar(stat = 'identity') +
  labs(x = 'Updates Per Day', y = 'Users')
```


### How many profiles do users have?
We'll use the data in [this look](https://looker.buffer.com/looks/3920) to answer this question. :) 

```{r}
profiles <- get_look(3920)
```

```{r include = FALSE}
colnames(profiles) <- c('user_id', 'profiles')
```

Let's summarize the profiles.

```{r}
summary(profiles$profiles)
```

Let's plot out the distribution.

```{r}
by_profile <- profiles %>%
  group_by(profiles) %>%
  summarise(users = n_distinct(user_id))

ggplot(by_profile, aes(x = as.factor(profiles), y = users)) +
  geom_bar(stat = 'identity') +
  labs(x = 'Number of Profiles', y = 'Users')
```

### How many users used share next in the past month.
Let's get these users with a SQL query.

```{r include = FALSE}
con <- redshift_connect()
```

```{sql connection=con, output.var=shared_next}
select
  a.user_id
  , count(a.*)
from actions_taken as a
left join users 
  on a.user_id = users.user_id
where a.full_scope like '%later next%'
and users.billing_plan = 'individual'
and a.date > (current_date - 30)
group by 1
```

Now let's join it into our original users

```{r}
# Join action counts
users <- users %>%
  left_join(shared_next, by = 'user_id')

# Replace NA's with 0s
users$count[is.na(users$count)] <- 0

# Indicate if they have used it
users <- users %>%
  mutate(has_shared_next = ifelse(count > 0, 'yes', 'no'))
```

Now let's see how many users and what percentage have shared next in the past 30 days.

```{r}
users %>%
  group_by(has_shared_next) %>%
  summarise(users = n_distinct(user_id)) %>%
  mutate(proportion = users/sum(users))
```

### Past 3 months?
We'll take the exact same approach, but adjust our query to look at users' actions in the past 90 days.

```{sql connection=con, output.var=shared_next}
select
  a.user_id
  , count(a.*)
from actions_taken as a
left join users 
  on a.user_id = users.user_id
where a.full_scope like '%later next%'
and users.billing_plan = 'individual'
and a.date > (current_date - 90)
group by 1
```

Now let's join it into our original users

```{r}
# Drop previous action count column
users$count <- NULL

# Join action counts
users <- users %>%
  left_join(shared_next, by = 'user_id')

# Replace NA's with 0s
users$count[is.na(users$count)] <- 0

# Indicate if they have used it
users <- users %>%
  mutate(has_shared_next = ifelse(count > 0, 'yes', 'no'))
```

Now let's see how many users and what percentage have shared next in the past 90 days.

```{r}
users %>%
  group_by(has_shared_next) %>%
  summarise(users = n_distinct(user_id)) %>%
  mutate(proportion = users/sum(users))
```

### How many users used a custom schedule in the past month.
Let's get these users with a SQL query.

```{r include = FALSE}
con <- redshift_connect()
```

```{sql connection=con, output.var=custom}
select
  a.user_id
  , count(a.*)
from actions_taken as a
left join users 
  on a.user_id = users.user_id
where a.full_scope like '%composer schedule%'
and users.billing_plan = 'individual'
and a.date > (current_date - 30)
group by 1
```

Now let's join it into our original users

```{r}
users$count <- NULL

# Join action counts
users <- users %>%
  left_join(custom, by = 'user_id')

# Replace NA's with 0s
users$count[is.na(users$count)] <- 0

# Indicate if they have used it
users <- users %>%
  mutate(has_custom = ifelse(count > 0, 'yes', 'no'))
```

Now let's see how many users and what percentage have shared next in the past 30 days.

```{r}
users %>%
  group_by(has_custom) %>%
  summarise(users = n_distinct(user_id)) %>%
  mutate(proportion = users/sum(users))
```

### Past 3 months

```{sql connection=con, output.var=custom}
select
  a.user_id
  , count(a.*)
from actions_taken as a
left join users 
  on a.user_id = users.user_id
where a.full_scope like '%composer schedule%'
and users.billing_plan = 'individual'
and a.date > (current_date - 90)
group by 1
```

Now let's join it into our original users

```{r}
users$count <- NULL

# Join action counts
users <- users %>%
  left_join(custom, by = 'user_id')

# Replace NA's with 0s
users$count[is.na(users$count)] <- 0

# Indicate if they have used it
users <- users %>%
  mutate(has_custom = ifelse(count > 0, 'yes', 'no'))
```

Now let's see how many users and what percentage have shared next in the past 30 days.

```{r}
users %>%
  group_by(has_custom) %>%
  summarise(users = n_distinct(user_id)) %>%
  mutate(proportion = users/sum(users))
```