---
title: "Custom Scheduled Updates"
output: github_document
---

## Motivation
The purpose of this analysis is to get an estimate of the percentage of Buffered posts that are shared now, shared to the queue, and scheduled at a custom time. It is in reference to [this issue](https://github.com/bufferapp/data-team/issues/63).

## Methodology
Because there are many users that share many updates through Buffer, we will take a small sample of users that shared updates in Buffer _in the past 7 days_. We will then determine the type of share by looking at the full scope of the `actions_taken` event associated with the update. Then we can run some summary statistics and visualize the breakdown.

## Data collection
It should be a fairly straightforward SQL query to get this data. Let's try it by querying `actions_taken`.

```{r include = FALSE}
library(dplyr); library(buffer); library(ggplot2)
```

```{r include = FALSE}
con <- redshift_connect()
```

```{sql connection=con, eval=FALSE}
select
  user_id
  , full_scope
  , count(distinct id)
from actions_taken
where full_scope like '%updates shared%'
and date > (current_date - 7)
group by user_id, full_scope
```

```{r}
# Save the data
# save(updates, file = "queue_updates.Rda")

# Load the data
load('queue_updates.Rda')
```

## Determine update types
Let's quickly examine the different values of `full_scope`.

```{r}
updates %>%
  group_by(full_scope) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

Hm. It seems like the last two words in the full scope in particular indicate whether the update was shared at a custom time, added to the queue, or shared now. 

If the `full_scope` contains "now", the update was shared now. Let's start with that.

```{r}
# Create new column
updates$type <- character(nrow(updates))

# Get index of updates that contain 'now' in full_scope
now_index <- grep("now", updates$full_scope, ignore.case = T)

# Set the type of these rows to 'now'
updates$type[now_index] = "now"
```

Cool, now let's do the same for the word "schedule"

```{r}
# Get index of updates that contain 'schedule' in full_scope
schedule_index <- grep("schedule", updates$full_scope, ignore.case = T)

# Set the type of these rows to 'custom_schedule'
updates$type[schedule_index] = "custom_schedule"
```

Now let's make an assumption: the rest of the updates were sent to the queue (in whichever slot).

```{r}
# Set remaining types to 'queue'
updates[updates$type == "", ]$type <- 'queue'
```

Great! I think we're ready now. 

## Exploratory analysis
Let's look at the overall percent of updates by type.

```{r}
# Group by update type
updates %>% 
  group_by(type) %>%
  summarise(users = n_distinct(user_id), updates = sum(count)) %>%
  mutate(updates_percent = updates / sum(updates)) %>%
  arrange(desc(updates_percent))
```

Cool, so around 51% of all updates sent in the past 7 days were sent to the queue, around 37% were sent via a custom schedule, and around 12% were sent now.

**_This might not tell the whole story however, as these percentages might be disproportionately influenced by users that send large numbers of updates._** It might be worth getting this breakdown for _each_ user, and then analyzing the distribution of that breakdown.

### Get breakdown per user
We want to group the updates by user and type of update. We first need to tidy the data a bit and make sure each user has a value for each type. Let's also only look at users that scheduled at least 5 updates

```{r}
# Group by user
by_user <- updates %>%
  group_by(user_id, type) %>%
  summarise(updates = sum(count)) %>%
  filter(updates >= 5)
``` 

Now we need to make sure the users have values for all three types of updats.

```{r}
library(tidyr)

# Spread the update types
spread <- by_user %>%
  spread(type, updates)

# Replace the new NAs with 0
spread[is.na(spread)] <- 0

# Now gather the updates into tidy format again
long <- spread %>%
  gather(type, updates, custom_schedule:queue)
```

There we go! Now for each user, we can calculate the proportion of updates for each type.

```{r}
# Calculate proportions for each user
proportions <- long %>%
  group_by(user_id, type) %>%
  summarise(updates = updates) %>%
  mutate(proportion = updates / sum(updates))
```

Now let's look at the plots of the distributions. 
```{r warning = FALSE, message = FALSE, echo = FALSE}
library(gridExtra)

# Custom plot
custom_plot <- ggplot(filter(proportions, type == 'custom_schedule'), aes(x = proportion)) + 
  geom_density() + 
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  labs(x = 'Proportion of Updates Custom Scheduled')

# Queue plot
queue_plot <- ggplot(filter(proportions, type == 'queue'), aes(x = proportion)) + 
  geom_density() + 
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  labs(x = 'Proportion of Updates Sent to Queue')

grid.arrange(custom_plot, queue_plot, nrow = 2)
```

Interesting, for both custom-scheduled posts and posts sent to the queue, users either did it a lot, or not very much. For custom-scheduled posts, most users only custom scheduled for a very small percentage of updates. However, some users sent a very high proportion of posts via a custom schedule.

For queue updates, it's a similar story, except _more_ users sent the vast majority of updates to the queue. :) 

## Conclusions
Most posts are sent directly to the queue, but a high proportion are also sent via a custom-schedule. The distribution doesn't seem to be even for users - either they send most directly to the queue or they send a low percentage of posts directly to the queue. 

Most users send a small percentage of their posts via a custom schedule, but there are also users that send a high proportion of updates via a custom schedule. :) 