---
title: "Small Business Churn Analysis"
output: github_document
---

## Motivation
The goal of this analysis is to predict the probability of a given subscription churning _on a given day_. This is a continuation of the [**subscription churn**](https://github.com/jwinternheimer/analysis/blob/master/Subscription_Churn.md) analysis done for all subscriptions created in 2016. 

We will use a similar approach in this analysis, looking at the same features and using the same methods, including decision trees, logistic regression, and random forests.

```{r setup, include = FALSE}
# Load libraries
library(buffer); library(dplyr); library(ggplot2); library(tidyr)
```

## Data collection
Let's grab the subscriptions data from Looker using the `buffer` package.

```{r}
# Get the data from Looker
subs <- get_look(3861)
```

```{r include = FALSE}
# Rename columns
colnames(subs) <- c('user_id', 'user_joined_at', 'team_members', 'profiles', 
                    'created_at', 'canceled_at', 'id', 'status', 'plan', 'billing_cycle', 
                    'amount', 'days_to_activate', 'country', 'referred_by_marketing', 
                    'visited_before_joining', 'charges', 'nps_responses', 'nps_score',
                    'helpscout_convos')

# Set dates as date objects
subs$user_joined_at <- as.Date(subs$user_joined_at, "%Y-%m-%d")
subs$created_at <- as.Date(subs$created_at, "%Y-%m-%d")
subs$canceled_at <- as.Date(subs$canceled_at, "%Y-%m-%d")

# Set NPS score as a factor
subs$nps_score <- as.factor(subs$nps_score)

# Set id as character type
subs$id <- as.character(subs$id)
subs$user_id <- as.character(subs$user_id)
```

We only want to look at subscriptions with the plan `business_v2_small_monthly` in this analysis, so let's filter the rest of the subscriptions out.

```{r}
# Filter out plans
small <- subs %>%
  filter(plan == 'business_v2_small_monthly')
```

Great, now we have 2756 small business v2 subscriptions to work with. Let's look at how many of them have churned.

```{r}
# Calculate subscription length
small <- small %>%
  mutate(length = as.numeric(canceled_at - created_at),
         churned = ifelse(is.na(canceled_at), 0, 1)) %>%
  mutate(churned_90_days = ifelse((length <= 90 & !(is.na(length))), 1, 0))

# See how many churned
small %>%
  group_by(churned) %>%
  summarise(subs = n_distinct(id)) %>%
  mutate(percent = subs / sum(subs))
```

Around 55% of these subscriptions churned. Now, let's gather the updates that these users sent **_each week after their subscription began_**. We'll do this using SQL.

```{r include = FALSE}
con <- redshift_connect()
```

```{sql eval = FALSE}
select 
  subscriptions.id
  , users.user_id 
  , date_trunc('week', updates.date) as week
  , count(distinct updates.id) as updates_count
from updates
left join users 
  on updates.user_id = users.user_id
left join stripe._subscriptions as subscriptions
  on users.billing_stripe_id = subscriptions.customer
where date(subscriptions.start) >= '2016-01-01'
and subscriptions.plan_id = 'business_v2_small_monthly'
and updates.status != 'service'
and updates.date >= subscriptions.start
and updates.date <= (subscriptions.start + 90)
group by 1, 2, 3
```

```{r include = FALSE}
# Save updates object
# save(updates, file = "~/business_updates.Rda")

# Load updates object
load('~/business_updates.Rda')
```

We also would like to look at the number of days that these users took an action in a given week.

```{sql eval = FALSE}
select 
  subscriptions.id
  , users.user_id 
  , date_trunc('week', actions_taken.date) as activity_week
  , count(distinct date(actions_taken.date)) as days_active
from actions_taken
left join users 
  on actions_taken.user_id = users.user_id
left join stripe._subscriptions as subscriptions
  on users.billing_stripe_id = subscriptions.customer
where date(subscriptions.start) >= '2016-01-01'
and subscriptions.plan_id = 'business_v2_small_monthly'
and actions_taken.date >= subscriptions.start
and actions_taken.date <= (subscriptions.start + 90)
group by 1, 2, 3
```

```{r include = FALSE}
# Save activity object
# save(activity, file = "~/business_activity.Rda")

# Load activity object
load('~/business_activity.Rda')
```

We can do the same with Helpscout conversations, but right now I'm not sure how to get the `user_id` and let's keep it simple for now.

## Data tidying
Let's start with a sequence of weeks.

```{r}
# Define function to set first day of week as Sunday
start_of_week <- function(date) {
  date - (setNames(c(6,0:5),0:6) [strftime(date,'%w')])
}

# Get a sequence of weeks
weeks <- seq(from = as.Date("2016-03-14"), to = Sys.Date(), by = "week")

# Make a dataframe
weeks <- as.data.frame(weeks) %>%
  mutate(week = start_of_week(weeks)) %>%
  select(-weeks)
```

We need to join the subscription data to the updates and activity datasets. 

```{r}
# Set as date object
updates$week <- as.Date(updates$week)
activity$activity_week <- as.Date(activity$activity_week)

# Join updates and activity
bus <- updates %>%
  right_join(weeks, by = "week") %>%
  full_join(activity, by = c("week" = "activity_week", "user_id" = "user_id", "id" = "id")) %>%
  left_join(select(small, c(1, 6, 7)), 
                   by = c("user_id" = "user_id", "id" = "id"))
```

Calculate churn week.

```{r}
# Churn week
bus <- bus %>% 
  mutate(churn_week = start_of_week(canceled_at)) %>%
  mutate(canceled_this_week = churn_week == week)
```



```{r}
# Determine if subscription churned that week
bus <- bus %>%
  mutate(churned_this_week = ifelse(week != churn_week | is.na(churn_week), 0, 1))
```

Let's also only look at the columns we need.

```{r}
# Select columns
bus <- bus %>%
  select(id:updates_count, created_at, canceled_at, churned:churn_week)
```

